---
title: "MachineLearning"
output: html_document
---

# Summary
This report describes the data analysis and the learn machine techniques taken to predict personal activities. in particular the goal of this analysis is to predict the manner in which people did an exercise. The data contains the measurements collected during group of people were training (e.g.data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants). The participants have been asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The data used to build the classifier was downloaded from [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv). After creating and evaluationg of the model, the classifier was tested on the follwoing [data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv) and delivered very promissing results, predicting all 20 test cases correctly.


In the following the R-code used for the analysis is presented.

## Data Loading & cleansing

```{r loading, results="hide"}
#load packages used
library(data.table)
library(caret)
#load training and test data
tr<-data.table(read.csv("pml-training.csv",sep=',',quote='"',na.strings = c("","NA")))
te<-data.table(read.csv("pml-testing.csv",sep=',',quote='"',na.strings = c("","NA")))
#load at the training data
str(tr)
summary(tr)
dim(tr)
#[1] 19622   160
```

## Data Cleaning:
After having a first look at the data there are obvious some errouneous rows and some empty columns.

```{r cleaning, results="hide"}
#remove the rows with "DIV/0"-error
tr_clean<-subset(tr[tr$new_window=='no',])
dim(tr_clean) #[1] 19216   160
summary(tr_clean)
#remove all columns having only NAs
tr_clean2<-tr_clean[,which(unlist(lapply(tr_clean, function(x)!all(is.na(x))))),with=F]
dim(tr_clean2)
summary(tr_clean2) #[1] 19216    60
#remove the first 6 columns, which do not have any measures but the name of the person, rows counts, date etc
tr_dt<-subset(tr_clean2,select=-c(1:6))
dim(tr_dt) #[1] 19216    54
```

## Data splitting: 
After an initial data cleaning is done I split the data into training and test data. The test data is used for the evaluation of the klassifier

```{r splitting}
#get 70% of the of the data for training and the rest 30% for evaluation 
tr_idx<-createDataPartition(tr_dt$classe,p=.7,list=FALSE)
trd<-subset(tr_dt[tr_idx[,1],])
ted<-subset(tr_dt[-tr_idx[,1],])
dim(trd) #[1] 13453    54
dim(ted) #[1] 5763   54
#check if the classes are also nearly equal distributed after the splitting
trd[,.N,by="classe"]
ted[,.N,by="classe"]
```

## Feature reduction:
We still have 54 features we can use to build a moder. To minimise the effort I am skipping the exploratory data analysis and using directly PCA to reduce the feature set 

```{r feature_selection}
#the 54th feature is skipped beacuse this is the one that need to be learned
preProc<-preProcess(trd[,-54,with=FALSE],method=c("pca"),thresh=0.95)
preProc #25 principal components are explaining 95 of the total variance of the data
trainPC<-predict(preProc,trd[,-54,with=FALSE])
```

## Model building: 
I am using Random Forest with Cross Validation of 10 folds to build a classifier
```{r model_building, cache=TRUE}
modelFitCV<-train(trd$classe~.,method="rf",data=trainPC,trControl = trainControl(method = "cv", number = 10))
testPC<-predict(preProc,ted[,-54,with=FALSE])
confusionMatrix(ted$classe,predict(modelFitCV,testPC))
``` 
With the model above it is achieved very high Sensitivity and Specificity in the same time on the training data. So I am using this model to make the prediction for this assignement.

## Prediction
Make the prediction for the submission

```{r prediction}
te_clean<-subset(te,select=colnames(tr_clean2)[-60])
te_clean2<-subset(te_clean,select=-c(1:6))
dim(te_clean2)
tePC<-predict(preProc,te_clean2)
res<-predict(modelFitCV,tePC)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(res)
```